---
title: "LSF array jobs with easypar"
author: "Giulio Caravagna"
date: "15 September 2019"
institute: "Institute for Cancer Research"
email: "giulio.caravagna@icr.ac.uk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{LSF array jobs with easypar}
  %\usepackage[UTF-8]{inputenc}
---

`easypar`  can be used  to submit jobs to the LSF cluster system. 

In particular, function  `run_lsf` generates all the required inputs to submit an
array of jobs of any size.

#### Requirements

To use  `run_lsf` you need:

-   your computation to be wrapped into a function `FUN` that can run as a stand-alone R application (so something that would run if called via `Rscript`);

- a input data.frame where every row rappresents an input for the function
(as for `apply` by row). 

The input should have column names without dots or spaces, as these
will be used to match the arguments of `FUN` with the same ordering of the function definition. So for instance
an input with 2-columns will only work if the function is expecting 2 parameters. The fist column will define the first parameter.

#### Output

This function generates 3 files:

- an R script wrapping the definition of your function `FUN`, as well as the required code to call that
function using parameters from the command line. Your function in this script is called with a fake name;

- a `tsv` file containing the input `PARAMS`, without any header (column names), and row names. 
function using parameters from the command line;

- a LSF array job submission script with `N` jobs where `N` are the rows of `PARAMS`.

#### Customisations (required)

Some basic customisations are possibles, and some are **required**. 

In particular, cluster-specific BSUB instructions have to be specified, as explained below, as well as other dependencies which depend on modules available on the cluster. 

Function `run_lsf` allows to:

- specify a list of modules that will be added as dependencies of the LSF submission script. 
  - For instance, `modules = 'R/3.4.0'` will generate the dependecy for a specific R version as `"module load R/3.4.0"` in the script.

- customize the BSUB parameters of the generated script. 
  
Default BSUB parameters parameters **have to be updated according to your cluster setup**.

```{r, eval=T}
library(easypar)

# Default parameters in the package
default_BSUB_config()
```

These are classical BSUB parameters:

- `-J`, the job ID;
- `-P`, the project ID;
- `-q`, the queue ID;
- `-n` and `-R`, the resources to allocate;
- `-W` the wall time of the jobs;
- `-o` and `-e`, the outputs and the error filenames. Notice that by default
we have the job array ID in the filename, so to have one log per job.

> It is required to modify the default values of `-P` and `-q`, the project and queue ID,
according to your LSF configuration. Otherwise, the submission script will 
generate an error becaue the default values do not mean anything.

Modifications are done to the default list of parameters; other BSUB flags can be
used as well. No checkings on their correctness are done by `easypar`.

```{r, eval=T}
custom_BSUB = default_BSUB_config()

# More informative job ID
custom_BSUB$`-J` = "bwa_aligner"

# A token for a project allowed to run on the cluster
custom_BSUB$`-P` = "DKSMWOP331"

# A queue name that is valid on the cluster
custom_BSUB$`-q` = "bioinformatics"
```

Once the BSUB has been customized, you can either:

- generate the submission scripts, and submit your job manually
- generate the submission scripts and submit the jobs with a `system` call. 

By default (`run = FALSE`) the `run_lsf` function outputs the shell command that should be used to submit the jobs, but leaves the user to submit the job. This is because we experienced some command line issues calling
modules with a system call.

An example computation follows.

```{r, eval=T}
# A simple function that prints some outputs
FUN = function(x, y){ 
  print(x, y) 
}

# input for 25 array jobs
PARAMS = data.frame(
  x = runif(25), 
  y = runif(25)
  )

# generates the input files, adding some module dependencies
run_lsf(FUN, PARAMS, 
        BSUB_config = custom_BSUB,
        modules = 'R/3.4.0'
        )
```

Command `bsub < EASYPAR_LSF_submission.sh` will submit the jobs to the cluster.
