---
title: "Vignette for easypar"
author: "Giulio Caravagna"
date: "15 November 2018"
institute: "Institute for Cancer Research"
email: "giulio.caravagna@icr.ac.uk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{Vignette for easypar}
  %\usepackage[UTF-8]{inputenc}
---

The `easypar` R package allows you to:

- run R functions in a parallel fashion in a trivial way.
- easily switch between parallel and serial executions of the same calls (at runtime).
- save results of paralle computation as far as they are produced (i.e., cache).

`easypar` can help if you have to run several different independent computations (for instance bootstrap estimates or multiple local-optimisations) and you want these to be parallel on multi-core architectures. `easypar` interfaces to `doParallel` in order to make this task easier to code, and to debug.

The idea is to exploit a code template and switch easily between parallel and sequential runs of a function. The code skeleton looks like this

```{r, eval=FALSE}
if(parallel)
{
   R = foreach(i = 1:N) %dopar% { ....fun....  }
}
else
{
   for(i in 1:N) { ....fun... }
}
```
where `f` is the actual computation. 

I want to use `parallel = FALSE` when I have to debug `f`, and eventually, I want to use `parallel = TRUE` to speed up computations. Parallel execution are hard to debug: inside `%dopar%`, tasks run in different memory spaces, and thus outputs (i.e., `print` etc) are asynchronous.

This piece of code is at the base of `easypar`, whose functioning is shown with some examples.

## Examples

Consider a dummy function `f` that sleeps for some random time and then print the output.

```{r}
f = function(x) 
{
  clock = 5 * runif(1)
  
  print(paste("Before sleep", x, " - siesta for ", clock))
  
  Sys.sleep(clock)
  
  print(paste("After sleep", x))
  
  return(x)
}
```
`f` runs as
```{r}
f(3)
```

**Input(s).** We want to run `f` on 4 inputs (random univariate numbers). We store them in a list where each position is a full set of parameters that we want to pass to each calls to `f` (list of lists), named according to the actual parameter names.


```{r}
inputs = lapply(runif(4), list)
print(inputs)
```

`easypar` provides a single function that takes as input `f`, its list of inputs and some execution parameters for the type of execution requested. The simplest call runs `f` in parallel, without seeing any output and just receiving the return values in a list as follows
```{r}
library(easypar)
easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE, outfile = NA)
```
We can control the amount (0 to 1) of cores to use at maximum (which are checked via `doPar`). Other combinations are also possible.

* make each thread dump to a shared `rds` file its result, implementing a cache which is usefull if one want to real-time analyze output results (with another process).
```{r, eval=TRUE}
easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE, outfile = NA, cache = "My_task.rds")

# Check
cache = readRDS("My_task.rds")
print(cache)
```

* get outputs to screen (asynchronous per thread) with `outfile`
```{r, eval=TRUE}
easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE, outfile = '')
```

* sequentially every tasks in a `for`-loop fashion
```{r, eval=TRUE}
easypar::run(FUN = f, PARAMS = inputs, parallel = FALSE, outfile = '')
```

## Runtime control of `easypar`

We can disable parallel executions easily.

We have a *global option* to force the execution to go serial, whatever its source code default behaviour is (`parallel = TRUE` will not work). 

When `f` is plugged in a tool and called as
```{r, eval=FALSE}
easypar::run(FUN = f, PARAMS = inputs)
```
which has default `parallel = TRUE`, and you set the global option `easypar.parallel`, `easypar` will run `f` sequentially. 
```{r}
options(easypar.parallel = FALSE)
easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE)
```

## Errors handling

```{r}
# Hopefully r will crash at least once but not all calls
f = function(x) 
{
  if(runif(1) > .5) stop("Boom!!")
  
  "Ok"
}

# Restore parallel and run
options(easypar.parallel = TRUE)
runs = easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE, outfile = NULL)

# inspect and filter function
numErrors(runs)
runs

filterErrors(runs)
```




