---
title: "Vignette for Easypar"
author: "Giulio Caravagna"
date: "15 November 2018"
output: html_document
---

A dummy function: sleeps for some random time and then print the output

```{r}
f = function(x) 
{
  clock = 5 * runif(1)
  
  print(paste("Before sleep", x, " - siesta for ", clock))
  
  Sys.sleep(clock)
  
  print(paste("After sleep", x))
  
  return(x)
}
```
which runs as
```{r}
f(3)
```

# Managing parallel/ serial tasks with easypar

Create 4 inputs (random univariate numbers), for which we want to compute our function `f`
```{r}
inputs = lapply(runif(4), list)
print(inputs)
```



We can run `f` in parallel, without seeing any output and just receiving the return values 
```{r}
library(easypar)
easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE, outfile = NA)
```

# Run in parallel, no output but cacheing computations 
easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE, outfile = NA, cache = "My_task.rds")

# check the cache
readRDS("My_task.rds")

# Run parallel, output to screen (asynchronous per thread)
easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE, outfile = '')

# Run sequentially, output to screen
easypar::run(FUN = f, PARAMS = inputs, parallel = FALSE, outfile = '')

# We have plugged in our tool a very complex function "f" which we call
# via easypar::run(FUN = f, PARAMS = inputs), but now we want to debug
# the runtime execution of f. The tool calls the execution with parallel
# set to TRUE, we just use the global option to force it to run without
# parallelism.
options(easypar.parallel = FALSE)
easypar::run(FUN = f, PARAMS = inputs, parallel = TRUE)



